Stability in human interaction networks: sector relative sizes, prominence of topological measures and time activity statistics.

### 2015-05-04 paper.tex
detalhe texto

\begin{quotation}
`The reason for the persistent plausibility of the typological approach, however, is not a static biological one, but just the opposite: dynamic and social. The fact that human society has been up to now divided into classes affects more than the external relations of men. The marks of social repression are left within the individual soul.' 
% `The conception of personality structure is the
%best safeguard against the
% inclination to attribute persistent trends in the
% individual to something
% "innate" or "basic" or "racial" within him. The
% Nazi allegation that natural, biological traits decide the total being of a % person
% would not have been such
% a successful political device
% had it not been possible to point to numerous
% instances of relative fixity in human behavior and to
% challenge those who
% thought to explain them on any basis other than a biological one.'
\emph{- Adorno et al, 1969, p. 747}
\end{quotation}

@@

Let a sector to which a vertex belongs yield a type for the corresponding participant. Accordingly, a participant may be considered peripheral, intermediary or a hub. This trivial consideration raises important questions: is this typology stigmatizing or prejudice inclined in any way? What else do we know about each type and about the typology itself?
These general considerations about this primitive typology will help us answer the first questions:
\begin{itemize}
	\item A participant belongs to many networks (his family, family of a friend, the email list to which he belongs, work friends, etc.).
	\item A participant might belong to all three sectors at the same time. Actually, it is reasonable to assume that almost every person belongs to all three sectors for some of the networks.
	\item The participant often transitions from one sector to another within a network.
	\item Reported stability of network structure is concomitant with the continuous change of participants and the type of each participant.
\end{itemize}

From these, we can assume that the typology is not stigmatizing because the type of an individual changes constantly~\cite{adorno}. That is to say, an individual is a hub in a number of network and peripheral in other networks, and even within a network he/she probably changes type along time.

Our approach to the second question was the inspection of network evolution from the raw data and using the tools for visualization~\cite{rcText,versinus}. This allowed us to infer relevant characteristics of each type, summarized as follows.

\begin{itemize}
    \item Core hubs usually have intermittent activity. Very stable activity was found on MET hubs, which motivated its integration to this work. There are reports in the literature of greater stability of participation in smaller communities~\cite{barabasiEvo}, which is the reason why the smaller number of participants in MET was considered coherent with the stable activity of hubs.
    \item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone. The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions. Therefore, intermediary vertices seem responsible for the network structure. For example, intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity.
    \item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs. These seem as authorities and contrast with participants that respond much more than receive responses.
    \item The most obvious community structure, as observed by a high clustering coefficient, is found only in peripheral and intermediary sectors.
\end{itemize}

This typology is a bridge from exact to human sciences and may be enriched with concepts from other typologies, such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.

@@

\section{Conclusions and further work}\label{sec:conc}
The characterization of interaction networks resulted from stability observations. 
Along temporal activity statistics, this work reports the stability of the formation of principal components and of the relative sizes of periphery, intermediary and hubs Erd\"os sectors.
These results suggests and eases the observance of types in both agents and the networks.

>>>>

\begin{quotation}
`The reason for the persistent plausibility of the typological approach, however, is not a static biological one, but just the opposite: dynamic and social.' 
% `The conception of personality structure is the
%best safeguard against the
% inclination to attribute persistent trends in the
% individual to something
% "innate" or "basic" or "racial" within him. The
% Nazi allegation that natural, biological traits decide the total being of a % person
% would not have been such
% a successful political device
% had it not been possible to point to numerous
% instances of relative fixity in human behavior and to
% challenge those who
% thought to explain them on any basis other than a biological one.'
\emph{- Adorno et al, 1969, p. 747}
\end{quotation}

@@

Let a sector to which a vertex belongs yield a type for the corresponding participant. Accordingly, a participant may be considered peripheral, intermediary or a hub. This trivial consideration raises important questions:  what do we know about each type and about the typology itself? Is this typology stigmatizing or prejudice inclined in any way?

These general considerations will help us answer such questions:
\begin{itemize}
	\item A participant belongs to many networks (his family, family of a friend, the email list to which he belongs, work friends, etc.).
	\item A participant might belong to all three sectors at the same time. Actually, it is reasonable to assume that almost every person belongs to all three sectors for some of the networks.
	\item The participant often transitions from one sector to another within a network.
	\item Reported stability of network structure is concomitant with the continuous change of participants and the type of each participant.
\end{itemize}

We can assume that the typology is not stigmatizing because the type of an individual changes constantly~\cite{adorno}. That is to say, an individual is a hub in a number of networks and peripheral in other networks, and even within a network he/she probably changes type along time.

To further insights we inspected raw data and visualized the evolving networks~\cite{rcText,versinus}. This allowed us to infer relevant characteristics of each type:

\begin{itemize}
    \item Core hubs usually have intermittent activity. Very stable activity was found on MET hubs, which motivated its integration to Supporting Information. There are reports in the literature of greater stability of activity in smaller communities~\cite{barabasiEvo}, which is the reason why the smaller number of participants in MET was considered coherent with the stable activity of MET hubs.
    \item Typically, the activity of hubs is trivial: they interact as much as possible, in every occasion with everyone. The activity of peripheral vertices also follows a simple pattern: they interact very rarely, in very few occasions. Therefore, intermediary vertices seem responsible for the network structure. Intermediary vertices may exhibit preferential communication to peripheral, intermediary, or hub vertices; can be marked by stable communication partners; can involve stable or intermittent patterns of activity.
    \item Some of the most active participants receive many responses with relative few messages sent, and rarely are top hubs. These seem as authorities and contrast with participants that respond much more than receive responses.
    \item The most obvious community structure, as observed by a high clustering coefficient, is found only in peripheral and intermediary sectors.
\end{itemize}

This typology bridges exact and human sciences and may be enriched with concepts from other typologies, such as Meyer-Briggs, Pavlov or the authoritarian types of the F-Scale~\cite{adorno}.

@@

\section{Conclusions and further work}\label{sec:conc}
The characterization of interaction networks resulted from stability observations. 
Along temporal activity statistics, this work reports the stability of the formation of principal components and of the relative sizes of periphery, intermediary and hubs Erd\"os sectors.
These results suggest and ease the observance of types in both agents and the networks.
